Notes on Control Plane Inference

### 2021-05-25 inference.tex
finish sketching proof

\paragraph{Commands}
Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$),
assumptions ($\assume b$), assertions ($\assert b$), table application
($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).

We define $\SKIP$ as syntactic sugar for $\assume \TRUE$.

@@

\begin{figure}[htp]
  %% \begin{minipage}[t]{0.5\textwidth}
  %% \[\begin{array}{r >{\triangleq}cl}
  %% \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  %% (h.f := e)\instr && h.f := e \\
  %% (\assume b)\instr && b \\
  %% (\assert b)\instr && \assert b\\
  %% (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  %% (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  %% (\{c\})\instr && \{c\instr\}\\
  %% (t.\apply())\instr && \\
  %% \multicolumn{3}{r}{
  %%   \qquad
  %%   \begin{array}[t]{l} \displaystyle
  %%     \{ \\
  %%     \quad \havoc \rho;\\
  %%     \quad \assume \rho \in t; \\
  %%     \quad \assume \rho.\matches = t.\keys; \\
  %%     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
  %%     \left(\begin{array}{l}
  %%       \assume \rho.\action = i;\\
  %%       t.\action[i](\rho.\data)
  %%     \end{array}\right)\\
  %%     \} \choice ( \\
  %%     \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
  %%     \quad \SKIP \\
  %%     )
  %%   \end{array}}
  %% \end{array}
  %% \]
  %% \end{minipage}\begin{minipage}[t]{0.45\textwidth}
\[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  (h.f := e)\instr && h.f := e \\
  (\assume b)\instr && \assume b \\
  (\assert b)\instr && \assert b\\
  (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  (t.\apply())\instr &&\\
  \multicolumn{3}{l}{
    \qquad
    \begin{array}[t]{l} \displaystyle
      \assume t.\matchrow(\rho_t); \\
      \assume \gamma_t.\keys = t.\keys; \\
    \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left(\begin{array}{l}
      \assume \rho_t.\action = i;\\
      \assume \gamma_t.\hitAction = \rho_t.\id;\\
      t.\action[i](\rho_t.\data)
    \end{array}\right)\\
  \end{array}}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation. For simplicity, we
    assume each action in a table has the same number (and types) of parameters}
  \label{fig:table-instrument}
\end{figure}

\begin{figure}[htp]
  \[\begin{array}{r cl}
  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
  \state\sigma_c &\triangleq
  & \displaystyle\bigwedge_{ t \in \tables(c) }
  \left(
  \begin{array}{>{\displaystyle}l}
  \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > |\sigma(t.\id)| \\
  \wedge \bigvee_{r_i \in \sigma(t.\id)} \left(\rho = r_i
  \wedge \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\right)\\
  \end{array}\right)
  \end{array}
  \]
  \caption{Symbolic State Encoding}
  \label{fig:table-instrument}
\end{figure}

@@

\begin{samepage}
\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  Given a set of variables $V \subseteq \Var$ that contains no ghost variables
  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$, well-formed state
  $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
  %% \[\denote{c}^\sigma\pkt =_{\fvs(c)} \denote{\assume\state\sigma_c;c\instr}\,\pkt.\]

  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc \vec{\rho_t}; \havoc \vec{\gamma_t}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
\end{theorem}
\end{samepage}

\begin{proof}
  Proceed by induction on $c$.
  \begin{enumerate}[align=left]
  \item[$(c = x:=e)$] Immediate.
  \item[$(c = \havoc x)$] Immediate.
  \item[$(c = \assume b)$] Immediate.
  \item[$(c = \assert b)$] Immediate.
  \item[$(c = c_1;c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
    follows:
    \[\begin{array}{ll}
    \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
    & \\
    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
    & \text{definition of }(\cdot)\instr \\
    = \denote{\havoc{\vec{\rho_t}}; \havoc{\vec\gamma_t}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
    & \text{LEMMAs INVOLVING SIGMA-DISTRIB \& ASSUME-DISTRIB} \\
    = \denote{\havoc{\vec{\rho_{t_1}}}\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
    & \text{by definition} \\
    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_1}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
    & \text{BY FREE-VARIABLE-ASSUMPTION COMMUTATIVITY LEMMA} \\
    = \denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}};\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
    & \text{By Commutativity of }\havoc\\
    = \denote{(\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr);
      (\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
    & \text{BY FREE-VARIABLE HAVOC COMMUTE LEMMA}\\
    = \denote{\havoc{\vec{\rho_{t_2}}}; \havoc{\vec\gamma_{t_2}};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
    \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
    & \text{BY DEFINITION} \\
    =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
    & \text{by IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1;c_2) \subseteq V \\
    =_V {\denote{c_2}^\sigma}^\dagger \left({\denote{c_1}^\sigma}^\dagger\pkt\right)
    & \text{by IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1;c_2) \subseteq V \\
    = \denote{c_1;c_2}^\sigma\pkt
    & \text{by definition of }\denote{-}^\sigma\\
    \end{array}\]

@@

  \item[$(c = c_1\choice c_2)$] Let $\vec{\rho_t}$ be the symbolic rows for tables $t
    \in \tables(c_1;c_2)$. We can write $\vec\rho_t$ as $\vec{\rho_t} = \vec{\rho_{t_1}} \vec{\rho_{t_2}}$
    where $\vec{\rho_{t_i}}$ is the set of symbolic rows in $c_i$. Let
    $\vec{\gamma_t}$ be the ghost variables rows for tables $t \in
    \tables(c_1;c_2)$. We can write $\vec{\gamma_t}$ as $\vec{\gamma_t} = \vec{\gamma_{t_1}} \vec{\gamma_{t_2}}$
    where $\vec{\gamma_{t_i}}$ is the set of ghost variables in $c_i$. Compute as
    follows:
    \[\begin{array}{ll}
    \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
    & \\
    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
    & \text{by definition of }(\cdot)\instr\\
    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
    & \text{BY LEMMA RELATING [] and STATE}\\
    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
    & \text{BY WEDGE-ASSUME LEMMA}\\
    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY PRODUCTS(;) AND SUMS([]) LEMMA} \\
    = \denote{\havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
      \choice \havoc{\vec{\rho_t}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr
      \choice \havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
    = \denote{\havoc{\vec{\rho_{t_1}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
    \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{by definition of }\denote{\cdot} \\
    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
    =_V \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt
    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
    = \denote{c_1 \choice c_2}^\sigma\pkt
    & \text{by definition of }\denote{\cdot}^\sigma
    \end{array}
    \]

@@

  \item[$(c = t.\apply())$] Let $\pkt$ and $\sigma$ be arbitrary. Let
    $[r_1,\ldots, r_n] = \sigma(t)$. There are two cases, corresponding to hit
    and miss.
    \begin{enumerate}
    \item[\textit{Case 1}.] Assume that there is some $1 \leq i \leq n$ such
      that $\pkt \models t.\matchrow(r_i)$. In fact, assume that $i$ is the
      smallest such $i$. Then, by definition, $\denote{t.\apply()}^\sigma\,\pkt
      = \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Show that the
      RHS is $V$-equal to this set. Let $P = \{\pkt\{\rho_t.* \mapsto
      v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$ \todo{clean this up}.
      Compute as follows:

>>>>

\paragraph{Commands}
Data plane programs $c \in \mathsf{Cmd}$ can be assignments ($h.f := e$), havoc statements ($\havoc x$),
assumptions ($\assume b$), assertions ($\assert b$), table application
($t.\apply()$), sequential composition ($c_1;c_2$ for two commands $c_1$ and $c_2$), and
nondeterministic choice ($c_1 \choice c_2$ for two commands $c_1$ and $c_2$).

We define $\SKIP$ as syntactic sugar for $\assume \TRUE$. We also lift havoc to
sets of variables $\havoc \{x_1,\ldots,x_n\} = \SKIP; \havoc{x_1}; \cdots; \havoc{x_n}$.

It is sensible to use sets here (rather than lists), because havocs commute:

\begin{lemma}[Havocs Commute]
  For any two variables $\{x_1,x_2\} \subseteq \Var$,
  \[\forall \sigma.\forall \pkt. \denote{\havoc{x_1};\havoc{x_2}}^\sigma\pkt = \denote{\havoc{x_2};\havoc{x_2}}^\sigma\pkt \]
\end{lemma}

\begin{proof}
  Let $x_1, x_2,\sigma,\pkt$ be given.
  Compute as follows:
  \[
  \begin{array}{ll}
    \denote{\havoc{x_1};\havoc{x_2}}^\sigma\pkt\\
    = {\denote{\havoc{x_2}}^\sigma}^\dagger\left(\denote{\havoc{x_1}}\right)^\sigma\pkt
    & \\
    = {\denote{\havoc{x_2}}^\sigma}^\dagger\{\pkt\{x_1 \mapsto v_1\} \mid v_1 \in \Value\}
    & \\
    = \{\pkt\{x_1 \mapsto v_1\}\{x_2 \mapsto v_2\} \mid v_1,v_2 \in \Value\}
    & \\
    = \{\pkt\{x_2 \mapsto v_2\}\{x_1 \mapsto v_1\} \mid v_1,v_2 \in \Value\}
    & \\
    = {\denote{\havoc{x_1}}^\sigma}^\dagger\{\pkt\{x_2 \mapsto v_2\} \mid v_2 \in \Value\}
    & \\
    = {\denote{\havoc{x_1}}^\sigma}^\dagger\left(\denote{\havoc{x_2}}^\sigma\pkt\right)\}
    & \\
    = \denote{\havoc{x_2};\havoc{x_2}}^\sigma\pkt
  \end{array}
  \]
\end{proof}

@@

\begin{figure}[htp]
  %% \begin{minipage}[t]{0.5\textwidth}
  %% \[\begin{array}{r >{\triangleq}cl}
  %% \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  %% (h.f := e)\instr && h.f := e \\
  %% (\assume b)\instr && b \\
  %% (\assert b)\instr && \assert b\\
  %% (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  %% (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  %% (\{c\})\instr && \{c\instr\}\\
  %% (t.\apply())\instr && \\
  %% \multicolumn{3}{r}{
  %%   \qquad
  %%   \begin{array}[t]{l} \displaystyle
  %%     \{ \\
  %%     \quad \havoc \rho;\\
  %%     \quad \assume \rho \in t; \\
  %%     \quad \assume \rho.\matches = t.\keys; \\
  %%     \quad \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
  %%     \left(\begin{array}{l}
  %%       \assume \rho.\action = i;\\
  %%       t.\action[i](\rho.\data)
  %%     \end{array}\right)\\
  %%     \} \choice ( \\
  %%     \quad \assume \forall \rho. \rho \in t \Rightarrow \rho.\matches \neq t.\keys;\\
  %%     \quad \SKIP \\
  %%     )
  %%   \end{array}}
  %% \end{array}
  %% \]
  %% \end{minipage}\begin{minipage}[t]{0.45\textwidth}
\[\begin{array}{r >{\triangleq}cl}
  \multicolumn{3}{l}{(\cdot)\instr : \Cmd \to \Cmd} \\
  (h.f := e)\instr && h.f := e \\
  (\assume b)\instr && \assume b \\
  (\assert b)\instr && \assert b\\
  (c_1;c_2)\instr && c_1\instr;c_2\instr\\
  (c_1 \choice c_2)\instr && c_1\instr \choice c_2\instr\\
  (t.\apply())\instr &&\\
  \multicolumn{3}{l}{
    \qquad
    \begin{array}[t]{l} \displaystyle
      \assume t.\matchrow(\rho_t); \\
      \assume \gamma_t.\keys = t.\keys; \\
    \displaystyle \bigchoice_{i=0}^{|t.\action| -1}
    \left(\begin{array}{l}
      \assume \gamma_t.\miss = 0;\\
      \assume \rho_t.\action = i;\\
      \assume \gamma_t.\hitAction = \rho_t.\id;\\
      t.\action[i](\rho_t.\data)
    \end{array}\right)\\
    \choice (\assume \gamma_t.\miss = 1; \SKIP)  \\
  \end{array}}
  \end{array}
  \]
  \caption{Symbolic Table instrumentation. For simplicity, we
    assume each action in a table has the same number (and types) of parameters}
  \label{fig:table-instrument}
\end{figure}

\begin{figure}[htp]
  \[\begin{array}{r cl}
  \multicolumn{3}{l}{\state\sigma_c : (\Value \to \Row^*) \to \BExpr} \\
  \state\sigma_c &\triangleq
  & \displaystyle\bigwedge_{ t \in \tables(c) }
  \left(
  \begin{array}{>{\displaystyle}l}
  \gamma_t.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > |\sigma(t.\id)| \\
  \wedge \bigvee_{r_i \in \sigma(t.\id)} \rho = r_i  \\
  \wedge \bigwedge_{r_i \in \sigma(t.\id)} \gamma_t.\hitAction = i \Leftrightarrow \rho_t = r_i \\
  \wedge \bigwedge_{r_i \in \sigma(t.\id)} \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1} \neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\\
  \end{array}\right)
  \end{array}
  \]
  \caption{Symbolic State Encoding}
  \label{fig:table-instrument}
\end{figure}

@@

\begin{samepage}
\begin{theorem}[Instrumentation Equivalence]
  \label{thm:instr-equiv}
  Given a set of variables $V \subseteq \Var$ that contains no ghost variables
  or symbolic variables. For every $c \in \Cmd$ such that $\fvs(c) \subseteq V$,
  well-formed state $\sigma : \Value \to \Row^*$, and packet $\pkt \in \Pkt$,
  where $R = \{\rho_t \mid t \in \tables(c)\}$, and $G = \{\gamma_t \mid t \in
  \tables(c)\}$,

  \[\denote{c}^\sigma\,\pkt =_V \denote{\havoc {R \cup G}; \assume{\state\sigma}_t; c\instr}\,\pkt  \]
\end{theorem}
\end{samepage}

\begin{proof}
  Proceed by induction on $c$.
  \begin{enumerate}[align=left]
  \item[$(c = x:=e)$] Immediate, since $R = \emptyset = G$, $\assume{\state\sigma}_{x:=e} = \TRUE$, and $(x:=e)\instr = (x:=e)$.
  \item[$(c = \havoc x)$] \textit{sim}.
  \item[$(c = \assume b)$] \textit{sim}.
  \item[$(c = \assert b)$] \textit{sim}.
  \item[$(c = c_1;c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
    as follows:
    \[\begin{array}{ll}
    \denote{\havoc {R \cup G}; \assume{\state\sigma_{c_1;c_2}};(c_1;c_2)\instr}\,\pkt
    & \\
    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1;c_2}}; c_1\instr;c_2\instr}\,\pkt
    & \text{definition of }(\cdot)\instr \\
    = \denote{\havoc{R\cup G}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
    & \text{LEMMAs INVOLVING SIGMA-DISTRIB \& ASSUME-DISTRIB} \\
    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2};\havoc{G_2}; \assume{\state\sigma_{c_1}};\assume{\state\sigma_{c_2}};c_1\instr;c_2\instr}\,\pkt
    & \text{by definition} \\
    = \denote{\havoc{R_1};\havoc{G_1}; \havoc{R_2}; \havoc{G_2}; \assume{\state\sigma_{c_1}};c_1\instr;\assume{\state\sigma_{c_2}};c_2\instr}\,\pkt
    & \text{BY FREE-VARIABLE-ASSUMPTION COMMUTATIVITY LEMMA} \\
    = \denote{(\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr);
      (\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr)}\,\pkt
    & \text{BY FREE-VARIABLE HAVOC COMMUTE LEMMA}\\
    = \denote{\havoc{R_2}; \havoc{G_2};\assume{\state\sigma_{c_2}};c_2\instr}^\dagger
    \left(\denote{\havoc{R_1}; \havoc{G_1}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
    & \text{BY DEFINITION} \\
    =_V {\denote{c_2}^\sigma}^\dagger \left(\denote{\havoc{\vec{\rho_{t_1}}}; \havoc{\vec\gamma_{t_1}}; \assume{\state\sigma_{c_1}};c_1\instr}\,\pkt\right)
    & \text{by IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1;c_2) \subseteq V \\
    =_V {\denote{c_2}^\sigma}^\dagger \left({\denote{c_1}^\sigma}^\dagger\pkt\right)
    & \text{by IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1;c_2) \subseteq V \\
    = \denote{c_1;c_2}^\sigma\pkt
    & \text{by definition of }\denote{-}^\sigma\\
    \end{array}\]

@@

  \item[$(c = c_1\choice c_2)$] Let $R_i = \{\rho_t \mid t \in \tables(c_i)\}$ for $i =
    1,2$. Notice that $R = R_1 \cup R_2$. Similarly, let $G_i = \{\gamma_t \mid
    t \in \tables(c_i)\}$ for $i = 1,2$. Notice that $G = G_1 \cup G_2$. Compute
    as follows:
    \[\begin{array}{ll}
    \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1 \choice c_2)\instr}\,\pkt\\
    & \\
    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1\choice c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt\\
    & \text{by definition of }(\cdot)\instr\\
    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1} \wedge \state\sigma_{c_2}}; (c_1\instr \choice c_2\instr)}\,\pkt
    & \text{BY LEMMA RELATING [] and STATE}\\
    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; (c_1\instr  \choice   c_2\instr)}\,\pkt
    & \text{BY WEDGE-ASSUME LEMMA}\\
    = \denote{\havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_1\instr
      \choice \havoc{R \cup G};\assume{\state\sigma_{c_1}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY PRODUCTS(;) AND SUMS([]) LEMMA} \\
    = \denote{\havoc{R \cup G}; \assume{\state\sigma_{c_1}}; c_1\instr
      \choice \havoc{R \cup G}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY UNUSED-VARIABLES -> TRIVIAL ASSUMPTIONS LEMMA} \\
    =_V \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr
      \choice \havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{BY UNUSEDVARIABLE HAVOC ELIMINATION LEMMA } \\
    = \denote{\havoc{R_1 \cup G_1}; \assume{\state\sigma_{c_1}}; c_1\instr}\,\pkt
    \cup \denote{\havoc{R_2 \cup G_2}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{by definition of }\denote{\cdot} \\
    =_V \denote{c_1}^\sigma\pkt \cup \denote{\havoc{\vec{\rho_{t_2}}};\havoc{\vec{\gamma_t}}; \assume{\state\sigma_{c_2}}; c_2\instr}\,\pkt
    & \text{By IH}(c_1) \text{ since } \fvs(c_1) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
    =_V \denote{c_1}^\sigma\pkt \cup \denote{c_2}^\sigma\pkt
    & \text{By IH}(c_2) \text{ since } \fvs(c_2) \subseteq \fvs(c_1 \choice c_2) \subseteq V \\
    = \denote{c_1 \choice c_2}^\sigma\pkt
    & \text{by definition of }\denote{\cdot}^\sigma
    \end{array}
    \]

@@

  \item[$(c = t.\apply())$] Let $\pkt$ and $\sigma$ be arbitrary. Let
    $[r_1,\ldots, r_n] = \sigma(t)$.

    Let $P = \{\pkt\{\rho_t.* \mapsto
      v\}\{\gamma_t.* \mapsto v' \mid v,v' \in \Value\} \}$ \todo{clean this up}.
      Compute as follows:

      \[\begin{array}{lcl}
      \denote{\havoc R \cup G; \assume {\state\sigma_{t.\apply()}}; t.\apply\instr}\,\pkt
      & \\
      = \denote{\assume {\state\sigma_{t.\apply()}}; t.\apply()\instr}^\dagger P
      & \text{by definition of }\denote{\cdot}^\sigma \\
      = \denote{t.\apply()\instr}^\dagger \{\pkt \mid \pkt \in P, \pkt \models \state\sigma_{t.\apply()} \}
      & \text{by definition of }\denote{\cdot}^\sigma \\
      = \denote{  \begin{array}{l} \displaystyle
          \assume \gamma_t.\keys = t.\keys; \\
          \displaystyle \bigchoice_{a=0}^{|t.\action| -1}
          \left(\begin{array}{l}
            \assume t.\matchrow(\rho_t); \\
            \assume \gamma_t.\miss = 0;\\
            \assume \rho_t.\action = a;\\
            \assume \gamma_t.\hitAction = \rho_t.\id;\\
            t.\action[a](\rho_t.\data)
          \end{array}\right) \\
          \choice (\assume \gamma_t.\miss = 1;\SKIP) \\
      \end{array}}
      \{\pkt \mid \pkt \in P, \pkt \models \state\sigma_{t.\apply()} \}
      & \text{by definition of }(\cdot)\instr \\
      = \displaystyle\bigcup_{a=0}^{|t.\action| - 1}
      \denote{t.\action[a](\rho_t.\data)}
      \left\{\pkt \in P\,\, \begin{array}{|ll}
        \pkt \models \state\sigma_{t.\apply()}, & \pkt \models t.\matchrow(\rho_t)\\
        \pkt \models \gamma_t.\keys = t.\keys, & \pkt \models \rho_t.\action = a \\
        \multicolumn{2}{|l}{\pkt \models \gamma_t.\hitAction = \rho_t.\id}
      \end{array}\right\}
      & \text{by definition of }\denote{\cdot}\\
      \qquad \cup\,\,\{\pkt \in P \mid \pkt \models \state\sigma_{t.\apply()} \wedge \gamma_t.\keys = t.\keys \wedge \gamma_t.\miss = 1 \}
      \end{array}\]

       There are two cases, corresponding to hit and miss. Let $M$ be the `miss
       set' i.e.
       \[M \triangleq \{\pkt \in P \mid \pkt \models \state\sigma_{t.\apply()} \wedge \gamma_t.\keys = t.\keys \wedge \gamma_t.\miss = 1 \},\]

       and let $H_a$ be the input `hit-set' for action $a$, i.e.
      \[ H_a \triangleq \left\{\pkt \in P\,\, \begin{array}{|ll}
        \pkt \models \state\sigma_{t.\apply()}, & \pkt \models t.\matchrow(\rho_t)\\
        \pkt \models \gamma_t.\keys = t.\keys, & \pkt \models \rho_t.\action = a \\
        \multicolumn{2}{|l}{\pkt \models \gamma_t.\hitAction = \rho_t.\id}
      \end{array}\right\}\]

    \begin{enumerate}
    \item[\textit{Case 1}.] Assume that there is some $1 \leq i \leq n$ such
      that $\pkt \models t.\matchrow(r_i)$. In fact, assume that $i$ is the
      smallest such $i$. Then, by definition, $\denote{t.\apply()}^\sigma\,\pkt
      = \denote{t.\action[r_i.\action](r_i.\data)}^\sigma\,\pkt$. Show that
      $M \cup \bigcup_{a=0}^{|t.\action|-1}\denote{t.\action[a](\rho_t.\data)} H_a$ is $V$-equal to this set.

      Show that $M = \emptyset$, by showing that the refinement is contradictory.

      Assume $\state\sigma_{t.\apply()}$, $\gamma_t.\keys = t.\keys$ and
      $\gamma_t.\miss = 1$. First expand the definition of
      $\state\sigma_{t.\apply()}$, which gives us $\gamma_t.\miss = 1
      \Leftrightarrow \gamma_t.\hitAction > n$, $\bigvee_{i}^n \rho = r_i$,
      $\bigwedge_{i=1}^n \gamma_t.\hitAction = i \Leftrightarrow \rho_t = r_i$,
      and

      \[\bigwedge_{i=1}^n \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys]\right)\]

      Since we know $\gamma_t.\keys = t.\keys$, we can equivalently write
      \[\bigwedge_{i=1}^n \left(\gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)\right)\]

      Since $\gamma_t.\miss = 1$ we know that $\gamma_t.\hitAction > n$, which
      means that for all $1 \leq j \leq n$, $\neg t.\matchrow(r_j)$, which
      contradicts the assumption that $t.\matchrow(r_j)$. Thus the refinement is
      contradictory and $M = \emptyset$. \hfill \checkmark

      Now, show that $\forall \pkt \in H_a. \rho_t = r_i$, by showing that the
      refinement implies that $\rho_t = r_i \wedge \bigwedge_{j \neq i}\rho_t \neq r_j$. We assume the following facts:
      \[ \state\sigma_{t.\apply()} \wedge \gamma_t = t.\keys \wedge \gamma.\hitAction = \rho_t.\id \wedge t.\matchrow(\rho_t) \wedge
      \rho_t.\action = a\]

      By unfolding the definition of $\state\sigma_{t.\apply()}$, we also know that
      $\gamma.\miss = 1 \Leftrightarrow \gamma_t.\hitAction > n$, $\bigvee_{i=1}^n \rho = r_i$, $\bigwedge_{i=1}^n \gamma_t.\hitAction = i \Rightarrow \rho_t = r_i$, and
      \[\bigwedge_{i=1}^n \left( \gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)[\gamma_t.\keys/t.\keys] \right)\]

      As before, since $\gamma_t.\keys = t.\keys$ we can write this as
      \[\bigwedge_{i=1}^n \left( \gamma_t.\hitAction \geq i \Leftrightarrow \bigwedge_{j=0}^{i-1}\neg t.\matchrow(r_j)\right)\]
